MATH 9999
\documentclass{article}
\usepackage{amstext,amssymb,amsmath,amsbsy}
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{color}
\usepackage{float}

\usepackage{hyperref}
\hypersetup{pdftex,colorlinks=false,allcolors=blue}
\usepackage{hypcap}

\input{_TMDD_Tex_Commands.tex}

\renewcommand{\vector}[2]{
\left[\begin{array}{c}
    {#1} \\
    {#2}
    \end{array}\right]
}
\renewcommand{\matrix}[4]{
\left[\begin{array}{cc}
    {#1} & {#2} \\
    {#3} & {#4}
    \end{array}\right]
}

\date{}

\begin{document}
{\Large{\bf Guiding dose selection of monoclonal antibodies using a new parameter (TFIRT) for characterizing ligand binding systems}} 
~
\\
\\
Abstract Word Count: XYZ
\\
Body Word Count (including figure captions): XYZ
\\
Number of Figures: XYZ 
\\
Number of Tables: XYZ
\\
Keywords: monoclonal antibody $|$ target mediated drug disposition $|$ pharmacokinetics and pharmacodynamics $|$ pharmacometrics

\bigskip

\def\thefootnote{\arabic{footnote}}
\begin{center}
    Sameed Ahmed\footnote{Departments of Mathematics, University of South Carolina, ahmed3@math.sc.edu},
    Miandra Ellis\footnote{School of Mathematical and Statistical Sciences, Arizona State University, mellis5@asu.edu},
    Hongshan Li\footnote{Department of Mathematics, Purdue University, li108@purdue.edu},
    Luca Pallucchini\footnote{Department of Mathematics, Temple University,  luca.pallucchini@temple.edu},
    Andrew M. Stein\footnote{Novartis Institute for BioMedical Research, 45 Sidney St, Cambridge MA, 02140 USA, andrew.stein@novartis.com}
\end{center}
Sameed, Miandra, Hongshan, and Luca all contributed equally to this manuscript.

\newpage

\tableofcontents

Figure are in a separate file at the end of this document.  They'll images themselves will need to be removed for CPT:PSP (06-Figures.tex). 

\section*{Abstract}
XYZ
\\
\\

\newpage
\input{01_Introduction.tex}
\input{02_Theory.tex}
\input{03_Methods_Results.tex}
\input{05_Discussion_Conclusion.tex}
\input{07_Appendix.tex}
\input{09_Appendix_SimilarityTransform.tex}

\section*{Study Highlights}
%150 word limit, 30 words of questions, so 180 word limit
\textbf{What is the current knowledge on the topic?}
XYZ
\\ \\
\textbf{What question did this study address?}
XYZ
\\ \\
\textbf{What this study adds to our knowledge?}
XYZ
\\ \\
\textbf{How might this change drug discovery, development and/or therapeutics?}
XYZ

\section*{Acknowledgements}
The authors would like to thank the IMA as the Math-to-Industry bootcamp was where this work began.  We'd also like to thank Ngartelbaye Guerngar who was also a member of the bootcamp team.  Finally, we'd like to thank Eduardo Sontag and Greg Thurber for many helpful discussions.

\section*{Conflict of Interest}
Andrew Stein is employed by Novartis Institute for BioMedical Research

\section*{Author Contributions}
XYZ analyzed the data, XYZ performed the research, XYZ wrote the manuscript.

\bibliography{_Biologics_Refs.bib} 
\bibliographystyle{ieeetr}

\input{08_Figures.tex}

\end{document}
