---
title: "Explore more realistic dosing scenarios"
author: "Andy Stein"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r, warning=FALSE}
suppressMessages(source("ams_initialize_script.R"))
suppressMessages(library(RxODE))
suppressMessages(library(dplyr))
suppressMessages(source("AFIRT_calculation.R"))
suppressMessages(source("ivsc_4cmtct_shedct.R"))
```

### Set initial parameters and functions
Here use doses ranging from 0.1-100 mg/kg or `r signif(.1*scale.mpk2nmol,2)`-`r signif(100*scale.mpk2nmol,2)` nmol because this is a more realistic dose range to consider. 

```{r}
dose.nmol.range = lseq(.1,100,7)*scale.mpk2nmol

model = ivsc_4cmtct_shedct()

tmax = 26*7 #days
tau  = 21   #days
compartment = 2

compare.theory.sim = function(model=model, param.as.double=param.as.double,
                          dose.nmol.range = dose.nmol.range,
                          tmax=tmax,tau=tau,dt=dt,compartment=compartment) {

  df_sim = data.frame() # put all simulations for different dose into one data frame
  for (dose.nmol in dose.nmol.range){
    row = lumped.parameters.simulation(model,
      param.as.double, dose.nmol, tmax, tau, compartment)
    df_sim = rbind(df_sim, row)
  }
  df_sim$Dose = dose.nmol.range
  
  df_thy = data.frame() # put all theoretical calculations of lumped parameters at different dose together
  for (dose.nmol in dose.nmol.range){
    row = lumped.parameters.theory(param.as.double=param.as.double, 
                                   dose.nmol=dose.nmol, 
                                   tau=tau)
    df_thy = rbind(df_thy, row)
  }
  df_thy = df_thy %>%
    mutate(Dose = dose.nmol.range,
           AFIRT = AFIRT.Kss)
  
  df_compare = bind_rows(df_thy,df_sim)
  df_compare = df_compare %>%
    arrange(Dose,type) %>%
    mutate_if(is.numeric,signif,2)
    

  #do a single simulation to make sure model is at steady state
  out = simulation(model=model, param.as.double=param.as.double, 
        dose.nmol=max(dose.nmol.range), tmax=tmax, tau=tau) %>%
    select(time,D1,D3,M3,DM3,Mfree.pct) %>%
    gather(cmt,value,-c(time))
  
  g = ggplot(out,aes(x=time,y=value,group=cmt,color=cmt))
  g = g + geom_line()
  g = g + scale.y.log10()
  print(g)
  
  #plot theory vs simulation
  df_plot = df_compare %>%
    select(-AFIRT.Kss,-AFIRT.Kd) %>%
    gather(param,value,-c(type,Dose))
  
  
  g = ggplot(df_plot, aes(Dose, value, group=type, color=type,linetype=type))
  g = g + scale.x.log10()
  g = g + scale.y.log10()
  g = g + geom_line()
  g = g + geom_point()
  g = g + facet_wrap(~param,scales = "free_y")
  print(g)

  return(df_compare)
}
```

### Herceptin
Because the receptor density of Herceptin is high, it make sense you need much higher doses to get good agreement.  In the clinic, Herceptin is dosed at only 6 mg/kg or `r 6*scale.mpk2nmol` nmol.  

What I believe is happening then is all the drug in the tumor (Cavg3) is getting sucked into the tumor and thus intratumoral concentrations are lower than the theory predicts. 

It would mkae sense to check this against other Herceptin modeling papers to see if they agree that there isn't saturation around 6 mg/kg.  Could check in with Greg Thurber at some point.

```{r, warning=FALSE}
d <- read_excel("../data/ModelF_Herceptin_Params.xlsx", 1)
param.as.double = d$Value
names(param.as.double) = d$Parameter

pdb.herceptin = param.as.double
p.herceptin = as.data.frame(t(param.as.double))
p.herceptin$drug = "Herceptin"

df_compare = compare.theory.sim(model = model,
                                param.as.double = param.as.double,
                                dose.nmol.range = dose.nmol.range,
                                tmax = tmax,
                                tau  = tau,
                                compartment = compartment)
#kable(df_compare)

```

### Atezolizumab
Atezolizumab is dosed at 1200 mg or `r 1200*scale.mg2nmol` nmol.  Andy doesn't currently know the kon and koff rates for atezolizumab.  He guessed koff = 3/d, but it could be 10 faster.

#### koff = 3/d
```{r, warning=FALSE}

d <- read_excel("../data/ModelF_Atezolizumab_Params.xlsx", 1)
param.as.double = d$Value
names(param.as.double) = d$Parameter

pdb.atezo = param.as.double
p.atezo = as.data.frame(t(param.as.double))
p.atezo$drug = "atezo"


df_compare = compare.theory.sim(model = model,
                                param.as.double = param.as.double,
                                dose.nmol.range = dose.nmol.range,
                                tmax = tmax,
                                tau  = tau,
                                compartment = compartment) 
#kable(df_compare)
```

#### koff = 30/d and kon scaled to keep Kd the same

Here, agreement between the simulation and theory is better

```{r warning=FALSE}

param.as.double[c("koff1","koff3")] = 30
param.as.double[c("kon1","kon3")] = 12

df_compare = compare.theory.sim(model = model,
                                param.as.double = param.as.double,
                                dose.nmol.range = dose.nmol.range,
                                tmax = tmax,
                                tau  = tau,
                                compartment = compartment) 

```